The binary liquid model for gravity-driven microchannel simulations

### 2010-11-08 Text/paper.tex
Multiple style and notation fixes in the whole paper.

\begin{abstract}
The motion towards exploration of nano- and microscale phenomena brings new
challenges to simulation tools. The continuous physics approach is limited on
those scales, and methods such as Molecular Dynamics (MD) or the Lattice
Boltzmann Method (LBM) become more appropriate. This work aims to
reproduce the gravity driven microchannel flow of a free surface liquid with the
help of the free-energy binary liquid lattice Boltzmann method. In the low capillary number
$\Ca$ regime, the flow is thoroughly described by Bretherton \cite{bretherton},
where the deposition film thickness is proportional to $\Ca^{2/3}$. The binary
liquid model we use is a continuous-interface multiphase model, which presents
the question of how to properly resolve the interface to obtain proper correlations \todo{between what?}.
In this work we investigate how the binary liquid model parameters and the grid resolution
need to be adjusted to properly simulate microchannel flows.  The presented results
should be of interest for industrial applications and to researchers simulating
Bretherton phenomena.
\end{abstract}

\section{Introduction}
Taylor/Bretherton \cite{bretherton} flow describes flow of long bubbles in
microchannels. It was found that the front meniscus film width is proportional
to $Ca^{2/3}$ for Capillary numbers less than $0.003$. While there are number
of works which simulate the Bretherton problem with the boundary value method
\cite{ingham-plates,heil-bretherton}, the simulation of Bretherton problem is of
difficulty for diffusion interface methods, where the interface is spread over
the few grid nodes. In this case the question of proper film width resolution
arises. While the discontinuous interface models are good to simulate
non-complicated geometries phenomena, the continuous interface methods are more
flexible to resolve problem evolving coalescence or droplet breakup. This work
concentrates on the resolving parameters range of the binary liquid lattice
Boltzmann method to validate it across the Bretherton problem.

Lattice Boltzmann method has emerged as successful method to simulate various
phenomena as hydrodynamics \cite{yu}, thermal-flows
\cite{karlin-minimalmodels}, micro-flows \cite{ansumali-small-knudsen},
ferrofluids \cite{kuzmin-aniso}, and multiphase flows
\cite{swift,Shan-chen:extended}. Based on kinetic nature and as a particle
method LBM tackles relatively easy complex geometries and the incorporation of
physics phenomena on the microscopic level, as multiphase models. Most
multiphase lattice Boltzmann models \cite{swift, Shan-chen:extended} resolve
the interface in the continuous way. Such representation as it was pointed out
brings issues with the film thickness resolution versus the interface
resolution. In other words, when it comes to the resolution of the film
thickness the diffusion interface should be resolved in such way as to have
negligible effect on physics in the film.

This work examines the binary liquids free-energy LBM model \cite{swift}. The
model simulates binary liquids within the assumption of the uniform overall
density. Therefore, the question of how to simulate the Bretherton problem
where two phases are completely of different densities (liquid and gas) is of
interest. However, the motion of a liquid slug and a bubble in a capillary is
governed by the capillary number. Therefore, the viscosities ratio not the
densities ratio is the critical parameter.

The goal of this paper is to give the answer as to how to choose related to the
binary liquid model parameters, viscosities ratio and grid
resolution to obtain the Bretherton/Taylor problem film thickness for the
diffusion interface model. The paper is organized as follows. We briefly
explain the binary liquid lattice Boltzmann model and summarize literature
review for the Bretherton problem. Then the parameters involved in the
simulation are examined and presented in the results section. The paper is
concluded with main findings.

\section{Bretherton bubble flow}
Bretherton in his significant work \cite{bretherton} studied a long bubble
moving in a tube filled with liquid. It was found that the deposition film is
proportional to $Ca^{2/3}$ in the range of small capillary numbers. Later on
it was figured \cite{wong-films,wong-pressure} that the film deposition
proportinal to $Ca^{2/3}$ only near the front meniscus and the film width is
variated through the bubble length for not infinite bubbles. Numerical
simulations \cite{giavedoni-numerical} and experimental studies
\cite{kreutzer-pressure-drop} show as well the deviation of rule $Ca^{2/3}$ for
capillary numbers more than $0.003$ and Reynolds number dependance. To be able
to predict consistently the flow pattern for capillaries in different range of
the parameters researchers validate their codes with the small capillary
numbers Bretherton problem.

There are a number of the numerical methods which
were used for the Taylor/Bretherton flow
simulation. \citet{vanbaten-circular} studied the mass transfer and film
thickness for rising bubbles in a circular capillar.
\citet{kreutzer-pressure-drop} performed numerical simulation by the
finite volume method the circular capillary for the number of different
Reynolds and capillary numbers. \citeauthor{wong-films} in
\cite{wong-films,wong-pressure} studied three-dimensional bubbles in the
polygonal capillaries and showed the derivations of different shapes in the
slug crossections and menisci appearance.
\citet{heil-bretherton,ingham-plates} studied the air finger propagation to
the two-dimensional channel for the range of Reynolds numbers and capillary
numbers. \citet{giavedoni-numerical} performed crossvalidation of the
finite element solution developed by them with the already published results
across the different capillary numbers. The solution is available for circular
and planar cases.

While the applicability of other methods are well established for the
simulation of the Bretherton/Taylor problem, the lattice Boltzmann
applicability and thorough parameters study to the best authors' knowledge is
not yet done. Especially that concerns the diffuse interface multiphase
models, where one needs to resolve the film accordingly. One should acknowledge
the works of
\citet{pagonabarraga-fingers} on meniscus in thin films for the fingering
phenomena. \citet{sehgal-microchannel} performed the lattice Boltzmann
simulations of two-dimensional channel flow for larger capillary numbers where
the authors found the discrepancies with the classical Bretherton theory having
limitations to the low capillary numbers regime \cite{giavedoni-numerical}.

While to resolve interface on a fine level for capillary numbers less than
$0.003$ is computationally extensive even for a two-dimensional case. For
instance for capillary number $0.01$ to properly resolve the interface one
needs around 22 GB memory requirement. Therefore, this work is based on the
comparison with the already established results which were in their case
validated over the Bretherton bubble flow. The work presents recipes for the
initialization and the optimal parameter range for the binary liquid LBM model,
which is described in the next section.

\section{Lattice Botlzmann binary liquid model}
The lattice Boltzmann equation operates on the rectangular grid representing the
physical domain. It utilizes the
probability distribution functions containing the information about the
macroscopic variables, such as density,
momentum. LBE consists from two parts: local collision and propagation operated
from one node to another in the certain
direction specified by the velocity set.  The lattice Boltzmann implementation
in a standard view representing the
collision and propagation steps:
\begin{equation}
\label{standard:implementation}
\begin{aligned}
&f_i^{*}(\bm{x},t)=\omega f_i^{eq}(\bm{x},t)-(1-\omega) f_i(\bm{x},t) +
F_i,&&\text{ collision step}\\
&f_i(\bm{x}+\bm{c_i},t+1)=f_i^{*}(\bm{x},t),&&\text{ propagation step}, 
\end{aligned}
\end{equation}
where $F_i$ is the external force population. The binary liquid LBM model is
based on the free-energy functional \cite{swift,landau} and operates with two
sets of populations one to track pressure and velocities fields, another one is
for the phase field. The model to be utilized is two-dimensional nine-velocity
D2Q9 model.

In terms of D2Q9 the equilibrium populations are as \cite{pooley-contact}:
\begin{equation}
\begin{aligned}
&f_i^{eq}=w_i 
\biggl(3
p_0 - k \phi \Delta \phi
+\frac{u_{\alpha}c_{i\alpha}}{c_s^2}+\frac{Q_{i\alpha\beta}u_{\alpha } u_ {
\beta}}{2 c_s^4}\biggr), i=1\div8\\
&f_0^{eq}=\rho-\sum_{i\neq0}{f_i^{eq}}\\
&g_i^{eq}=w_i(\Gamma \mu + \frac{\phi c_{i\alpha} u_{i\alpha}}{c_s^2}+\phi^m
\frac{Q_{i\alpha\beta}u_{\alpha}u_{\beta}}{2 c_s^4}), i=1\div8 \\
&g_0^{eq}=\phi-\sum_{i\neq0}{g_i^{eq}}\quad,
\end{aligned}
\end{equation}
where $\Gamma$ is the mobility parameter; chemical potential
$\mu=-A\phi+A\phi^3-k\Delta\phi$; $k$ is related to the surface
tension; $A$ is the parameter of the free-energy model; bulk pressure
$p_0=c_s^2 \rho +A (-0.5 \phi^2+0.75 \phi^4)$. D2Q9 related parameters are
$w_i=\left\{\frac{4}{9},\frac{1}{9},\frac{1}{9},\frac{1}{9},\frac{1}{9},
\frac{1}{36},\frac{1}{36},\frac{1}{36},\frac{1}{36}\right\}$, and
$Q_{i\alpha\beta}=c_{i\alpha} c_{i\beta} - c_s^2 \delta_{\alpha\beta}$ with
the sound speed parameter $c_s^2=\dfrac{1}{3}$. Weights related to the
inclusion of the surface tension coefficient are as follows:
$w^{xx}_{1-2}=w^{yy}_{3-4}=1/3$, $w^{xx}_{3-4}=w^{yy}_{1-2}=-1/6$,
$w^{xx}_{5-8}=w^{yy}_{5-8}=-1/24$, $w^{xy}_{1-4}=0$, $w^{xy}_{5-6}=1/4$ and
$w^{xy}_{7-8}=-1/4$. Note that the system restores the macroscopic
equations as:
\begin{equation}
\begin{aligned}
&\partial_t \rho+ \partial_{\alpha} \rho u_{\alpha}=0\\
&\rho\left(\partial_t+u_{\beta}\partial_{\beta}\right) u_{\alpha}=
-\partial_{\alpha}P_{\alpha \beta} +
\nu\partial_{\beta}\left(\partial_{\alpha}u_{\beta}+\partial_{\beta} u_{\alpha}
+ \frac{1}{3}\partial_{\gamma} u_{\gamma} \delta_{\alpha\beta}\right)\\
&\partial_t \phi + \partial_{\alpha} \phi u_{\alpha}=M \Delta \mu,
\end{aligned}
\label{binary:fluid:system}
\end{equation}
where the viscosity $\nu=c_s^2 (\tau-1/2)$ and mobility parameter
$M=\Gamma(\tau_{\phi}-1/2)$. The system allows the separation of the liquid
phase with $\phi=1$ and so-called gas phase with $\phi=-1$. Note that the
spatial variation of the relaxation parameter depends on the relaxation
parameter of $\tau_{gas}$ and $\tau_{liq}$ linearly as
$\tau=\tau_{gas}+\frac{\phi+1}{2}\tau_{liq}$. Note that the surface tension in
the framework of binary liquid model as $\sqrt{8 k A}{9}$.

\section{Lattice Botlzmann benchmark}
To properly simulate and to compare simulations results with the already
performed simulations one needs to design a numerical benchmark by addressing
the following challenges for the lattice Boltzmann method:
\begin{itemize}
 \item The diffusive interface nature of the multiphase model should not
influence the film width results. This is obviously achieved only for certain
resolutions of the diffuse interface in comparison with the film width.
 \item The spurious currents arising from the surface tension discretization
can influence the mass transfer in the thin film therefore comprimising the
overall solution. Therefore the bubble should be long enough to decrease the
influence of the spurious currents.
 \item The wettability coefficient which is in principle can control the
dynamic angle and menisci appearance \cite{pagonabarraga-finger} should not
influence on the overall solution.
 \item The free-energy model examined in the work is one density/different
viscosities model. Classical Bretherton problem is the free-interface problem
where the film thickness is established by itself. The viscosities ratio should
be large as it is indicated in \cite{pagonabarraga-fingers}, however no studies
address this issue.
  \item The Bretherton problem addresses the bubbles flow driven by pressure
difference but not body force. However, it's relatively difficult to impose the
pressure difference in the framework of the binary liquid model because two
boundary conditions with the associated gradients and laplacians are required.
The pressure boundary conditions for the binary liquid model are currently
under development. Preliminary results show that the differences between
pressure difference flows and body force driven flows are small for the film
deposition thickness, the interest of this study. Thus, the work's focus is at
the body force driven flows because of the better simulation stability and
realization along with the periodic boundary conditions.
 \item Most simulations are done across the circular shape of channel,
which is quite difficult to address in terms of lattice Boltzmann free energy
framework where one needs to calculate the laplacians near the boundaries.
Although one can apply certain finite difference stencils
\cite{arnold-boundary,hunt-boundary} developed for the curved boundaries
however the analysis of error which can be addressed to the boundary conditions
treatment becomes an issue. Therefore the benchmark is for an axis-symmetric
case. 
\end{itemize}
Given all the concerns and challenges the suggested Lattice Boltzmann framework
is the two-dimensional flow driven by a body force. However, the analytical
solution to this problem, to the best authors' knowledge, is not yet known.
Though some authors \cite{sehgal-microchannel} use the gravity driven model to
validate the Bretherton film thickness, the latter should be used with caution
since the velocity of the bubble is different from the fluid velocity and a
bubble shape
is different in front and rear menisci. \citet{wong-films} indicated that the
film width for polygonal capillaries
changes from $Ca^{1/2}$ on distances $x\propto Ca^{-1}$ to $Ca^{2/3}$ for
distances $x \gg Ca^{-1}$. Therefore, for a bubble to have a formed shape in
the  front meniscus we picked up the bubble length as $5$ channel heights. The
channel length is taken to minimize the influence of one bubble on another,
because of the periodicity of boundary conditions. For the simulations the
channel length is taken as $3$ bubble lenghts giving all together $15$ channel
heights. Figure \ref{fig:benchmark:sketch} represents the benchmark sketch.
\begin{figure}
%\includegraphics[width=0.97\textwidth]{}
\caption{{\color{red} Ask Sasha to help here. The benchmark sketch.}
\label{fig:benchmark:sketch}}
\end{figure}

\section{Results}
\subsection{The nondimensionalization and initialization procedure}
This section is to introduce the procedure of nondimensionalization and
initialization techniques to be able to predict parameters necessary for
simulation. The capillary number governs the interface thickness and defined as:
\begin{equation}
Ca=\frac{U_{bubble} \mu_{liq}}{\gamma},
\end{equation}
where $\mu_{liq}$ is the liquid viscosity, $\gamma$ is the surface tension. The
bubble velocity $U_{bubble}$ is taken in the center of the bubble where the film
thickness is examined. 
While parameters as $\mu_{liq}$, $\gamma$ and capillary number $Ca$ are given
or defined through the model parameters, the velocity $U_{bubble}$ is not
defined. In the simulation we approximately match the velocity with the
Poiseuille profile center velocit, which is given by the equation:
\begin{equation}
\begin{aligned}
U_{bubble} \approx U_{liq}=\frac{H^2}{8
\mu_{liq}}\frac{\mathrm{d}P}{\mathrm{d}x}\\
U_{bubble}\approx
\frac{{N_y}^2}{\frac{8}{6}(\tau_{liq}-\frac{1}{2})}\frac{\mathrm{d}P}{\mathrm{d}
x } ,
\end{aligned}
\end{equation}
By taking the Poiseuille profile one can estimate the body force through the
capillary number and grid resolution:
\begin{equation}
\label{poiseuille:velocity:center}
\begin{aligned}
&\frac{\mathrm{d}P}{\mathrm{d}x}=\frac{8}{{N_y}^2}\gamma Ca.
\end{aligned}
\end{equation}
In reality we iterate the body force value
until the desired velocity and the capillary number as well are achieved.

After one resolves necessary parameters to run a simulation, one needs to
initialize the macroscopic fields and populations.
The velocity is initialized with $0$ everywhere. Populations are initialized
with the equilbrium values for the binary liquid model including all the phase
gradients and laplacians. The liquid phase is initialized with $1$ and the
gas phase is initialized with $-1$. Though the initialization film thickness
is not involved in the final result for the front meniscus, Fig.
\ref{fig:different:initialization:widths}, we keep the initialized film
thickness as close as possible to the already obtained numerical film
thicknesses \cite{giavedoni-numerical}. Otherwise, the simulation time for the
bubble to obtain a steady-state form increases.
\begin{figure}
\includegraphics*[bb=-20 360 620 430,width=0.97\textwidth]
{Figures/Init/initbegin12.eps}
\includegraphics*[bb=-20 360 620 430,
width=0.97\textwidth]{Figures/Init/initbegin16.eps}
\includegraphics*[bb=-20 360 620 430,
width=0.97\textwidth]{Figures/Init/initbegin20.eps}
\includegraphics*[bb=-20 360 620
430,width=0.97\textwidth]{Figures/Init/initfinish12.eps}\\
\includegraphics*[bb=-20 360 620
430,width=0.97\textwidth]{Figures/Init/initfinish16.eps}\\
\includegraphics*[bb=-20 360 620
430,width=0.97\textwidth]{Figures/Init/initfinish20.eps}\\
\caption{The phase plots for different bubble widths as
$H_{eff}-12$,$H_{eff}-16$ and
$H_{eff}-20$. The grid for the simulation is $102\mathrm{x}1501$. The phase
profiles were obtained after $200000$ time steps. One can see that even the
system is initialized with different bubble volumes the system comes to the
shape, where the film thickness is conserved. For the given parameters the
interface thickness is $0.0615$, $0.0611$, $0.0605$.
\label{fig:different:initialization:widths}}
\end{figure}

>>>>

\begin{abstract}
The motion towards exploration of nano- and microscale phenomena brings new
challenges to simulation tools. The continuous physics approach is limited on
those scales, and methods such as Molecular Dynamics (MD) or the Lattice
Boltzmann Method (LBM) become more appropriate. This work aims to
reproduce the gravity driven microchannel flow of a free surface liquid with the
help of the free-energy binary liquid lattice Boltzmann method. In the low capillary number
$\Ca$ regime, the flow is thoroughly described by Bretherton \cite{bretherton},
with the deposition film thickness proportional to $\Ca^{2/3}$. The binary
liquid model we use is a continuous-interface multiphase model, which presents
the question of how to properly resolve the interface to obtain proper correlations \todo{between what?}.
In this work we investigate how the binary liquid model parameters and the grid resolution
need to be adjusted to properly simulate microchannel flows.  The presented results
should be of interest for industrial applications and to researchers simulating
Bretherton phenomena.
\end{abstract}

\section{Introduction}
The Taylor/Bretherton \cite{bretherton} flow describes the flow of long bubbles in
microchannels. It was found that the front meniscus film width is proportional
to $\Ca^{2/3}$ for Capillary numbers less than $0.003$. While there are number
of works which simulate the Bretherton problem with the boundary value method
\cite{ingham-plates,heil-bretherton}, its simulation using diffusive interface methods
is difficult.  If the interfaec is spread over several grid nodes, the question of
proper film width resolution arises. While the discontinuous interface models
are good at simulating pheonomena in simple geometries, the continuous interface
methods are more flexible and capable of resolving problems involving coalescence
or droplet breakup. This work
concentrates on resolving the parameter range of the binary liquid lattice
Boltzmann method to validate it the Bretherton regime.

The lattice Boltzmann method has emerged as a successful method to simulate
a wide variety of phenomena including hydrodynamics \cite{yu}, thermal flows
\cite{karlin-minimalmodels}, microflows \cite{ansumali-small-knudsen},
ferrofluids \cite{kuzmin-aniso}, and multiphase flows
\cite{swift,Shan-chen:extended}. Thanks to its kinetic nature, LBM as a particle
method easily tackles complex geometries and the incorporation of
physical phenomena on the microscopic level, as is the case in multiphase models. Most
multiphase lattice Boltzmann models \cite{swift, Shan-chen:extended} resolve
the interface in a continuous way. Such a representation
brings issues with the film thickness resolution versus the interface
resolution. In other words, when it comes to the resolution of the film
thickness the diffusion \todo{diffusive?} interface should be resolved in such a way as to have
negligible effect on the physics of the film.

This work examines the binary liquid free-energy LB model \cite{swift}. The
model simulates two liquids with the assumption of uniform overall
density. Therefore, the question of how to simulate the Bretherton problem
where the two phases have two completely different densities (i.e. the two fluids
are liquid and gas) is of interest. However, in the case of the motion of a
liquid slug and a bubble in a capillary the governing factor is the capillary number
and the ratio of the viscosities, not of the densities, is the critical parameter.

The goal of this paper is to give an answer as to how to choose the
binary liquid model parameters, the viscosities ratio and the grid
resolution, in order to obtain correct Bretherton/Taylor film thickness.
The paper is organized as follows.  First, we briefly
explain the binary liquid lattice Boltzmann model and summarize the literature
for the Bretherton problem. Then, the parameters involved in the
simulation are examined and presented in the results section. The paper is
concluded with a summary of the main findings.

\section{Bretherton bubble flow}
Bretherton in his significant work \cite{bretherton} studied a long bubble
moving in a tube filled with liquid. It was found that the deposition film width
is proportional to $\Ca^{2/3}$ in the range of small capillary numbers. Later on
it was realized \cite{wong-films,wong-pressure} that the film deposition
is proportional to $\Ca^{2/3}$ only near the front meniscus and that the film width
varies throughout the bubble length for bubbles of finite length. Numerical
simulations \cite{giavedoni-numerical} and experimental studies
\cite{kreutzer-pressure-drop} showed as well a Reynolds number dependence, and
a deviation from the $\Ca^{2/3}$ rule for capillary numbers larger than $0.003$.
To be able to predict consistently the flow pattern for capillaries in
different ranges of parameters, simulations are developed and their codes
are typically validated with the small capillary numbers Bretherton problem.

There are a number of numerical methods which
were used for the simulation of the Taylor/Bretherton flow.
\citet{vanbaten-circular} studied the mass transfer and film
thickness for rising bubbles in a circular capillary.
\citet{kreutzer-pressure-drop} used the finite volume method to perform
simulations of a circular capillary for a number of different
Reynolds and capillary numbers. \citeauthor{wong-films} in
\cite{wong-films,wong-pressure} studied three-dimensional bubbles in
polygonal capillaries and showed the derivations of different shapes in the
slug crossections and menisci appearance.
\citet{heil-bretherton,ingham-plates} studied air finger propagation in
a two-dimensional channel for a range of Reynolds and capillary
numbers. \citet{giavedoni-numerical} performed crossvalidation of the
finite element solution developed by them with previously published results.
The solution is available for the circular and planar cases.

While the applicability of other methods is well established for the
simulation of the Bretherton/Taylor problem, this is not the case for
the lattice Boltzmann method. A thorough parameter study for this approach
has not yet been done to the best of the authors' knowledge.

One should acknowledge the works of \citet{pagonabarraga-fingers} on menisci
in thin films for fingering phenomena. \citet{sehgal-microchannel} performed lattice Boltzmann
simulations of two-dimensional channel flows for larger capillary numbers, and
the authors found discrepancies with the classical Bretherton theory, which
is limited to the low capillary number regime \cite{giavedoni-numerical}.

To resolve the interface on a fine level for capillary numbers less than
$0.003$ is computationally intensive even in the two-dimensional case
(the amount of memory necessary to perform the simulation is on the order
of tens or hundreds of gigabytes).  This work is therefore based on a
comparison with the already established results, which were themselves
validated for the Bretherton bubble flow.  In this paper, we present
recipes for the initialization of the simulations and discuss optimal
parameter ranges for the binary liquid LB model.

\section{Lattice Boltzmann binary liquid model}
The lattice Boltzmann equation (LBE) operates on a rectangular grid representing the
physical domain. It utilizes
probability distribution functions (also known as particle populations)
containing information about
macroscopic variables, such as fluid density and momentum. LBE consists of
two parts: a local collision step, and a propagation step which transports
information from one node to another in certain
directions specified by the discretized velocity set.
The LBE is typically implementated as follows:
\begin{equation}
\label{standard:implementation}
\begin{aligned}
&f_i^{*}(\bm{x},t)=\omega f_i^{eq}(\bm{x},t)-(1-\omega) f_i(\bm{x},t) +
F_i,&&\text{ collision step}\\
&f_i(\bm{x}+\bm{c_i},t+1)=f_i^{*}(\bm{x},t),&&\text{ propagation step}, 
\end{aligned}
\end{equation}
where $F_i$ is the external force population. The binary fluid LB model is
based on a free-energy functional \cite{swift,landau}, and operates with two
sets of populations: one to track the pressure and velocity fields, and the
other to represent the phase field.

The model we use in this work is a two-dimensional nine-velocity (D2Q9) model,
with equilibrium populations \cite{pooley-contact}:
\begin{equation}
\begin{aligned}
&f_i^{eq}=w_i 
\biggl(3
p_0 - k \phi \Delta \phi
+\frac{u_{\alpha}c_{i\alpha}}{c_s^2}+\frac{Q_{i\alpha\beta}u_{\alpha } u_ {
\beta}}{2 c_s^4}\biggr), i=1\div8\\
&f_0^{eq}=\rho-\sum_{i\neq0}{f_i^{eq}}\\
&g_i^{eq}=w_i\left(\Gamma \mu + \frac{\phi c_{i\alpha} u_{i\alpha}}{c_s^2}+\phi^m
\frac{Q_{i\alpha\beta}u_{\alpha}u_{\beta}}{2 c_s^4}\right), i=1\div8 \\
&g_0^{eq}=\phi-\sum_{i\neq0}{g_i^{eq}}\quad,
\end{aligned}
\end{equation}
where $\Gamma$ is the mobility parameter; the chemical potential
$\mu=-A\phi+A\phi^3-k\Delta\phi$; $k$ is related to the surface
tension; $A$ is a parameter of the free-energy model  The bulk pressure
is expressed as $p_0=c_s^2 \rho +A (-0.5 \phi^2+0.75 \phi^4)$. 
Parameters speficic to the D2Q9 grid are the weights
$w_i=\left\{\frac{4}{9},\frac{1}{9},\frac{1}{9},\frac{1}{9},\frac{1}{9},
\frac{1}{36},\frac{1}{36},\frac{1}{36},\frac{1}{36}\right\}$, and the tensor
$Q_{i\alpha\beta}=c_{i\alpha} c_{i\beta} - c_s^2 \delta_{\alpha\beta}$ with
the sound speed parameter $c_s^2=1/3$.  The weights related to the
inclusion of the surface tension coefficient into the equations are as follows:
$w^{xx}_{1-2}=w^{yy}_{3-4}=1/3$, $w^{xx}_{3-4}=w^{yy}_{1-2}=-1/6$,
$w^{xx}_{5-8}=w^{yy}_{5-8}=-1/24$, $w^{xy}_{1-4}=0$, $w^{xy}_{5-6}=1/4$ and
$w^{xy}_{7-8}=-1/4$. The system can be shown to restore the macroscopic
fluid equations as:
\begin{equation}
\begin{aligned}
&\partial_t \rho+ \partial_{\alpha} \rho u_{\alpha}=0\\
&\rho\left(\partial_t+u_{\beta}\partial_{\beta}\right) u_{\alpha}=
-\partial_{\alpha}P_{\alpha \beta} +
\nu\partial_{\beta}\left(\partial_{\alpha}u_{\beta}+\partial_{\beta} u_{\alpha}
+ \frac{1}{3}\partial_{\gamma} u_{\gamma} \delta_{\alpha\beta}\right)\\
&\partial_t \phi + \partial_{\alpha} \phi u_{\alpha}=M \Delta \mu,
\end{aligned}
\label{binary:fluid:system}
\end{equation}
where $\nu=c_s^2 (\tau-1/2)$ is the viscosity and
$M=\Gamma(\tau_{\phi}-1/2)$ is the mobility parameter. The system allows the separation of the liquid
phase with $\phi=1$ and a so-called gas phase with $\phi=-1$. The
relaxation time is taken to depend linearly on the relaxation
times $\tau_{\mathrm{gas}}$ and $\tau_{\mathrm{liq}}$:
$\tau=\tau_{\mathrm{gas}}+\frac{\phi+1}{2}\tau_{\mathrm{liq}}$.
The surface tension in the framework of the binary liquid model is $\sqrt{8 k A}{9}$ \todo{is this right?  should the 9 be in the
denominator?  is not, perhaps put before the root?}.

\section{Lattice Boltzmann benchmark}
To properly simulate and to compare simulation results with the data
published in the literature, one needs to design a numerical benchmark addressing
the following challenges for the lattice Boltzmann method:
\begin{itemize}
 \item The diffusive nature of the interface in the multiphase model should not
influence the film width results. This is obviously achieved only for certain
resolutions of the interface in comparison with the film width.
 \item Spurious currents arising from surface tension discretization
can influence mass transfer in the thin film, thereby comoromising the
overall solution. The bubble should therefore be long enough to decrease the
influence of spurious currents.
 \item The wettability coefficient which in principle can control the
dynamic angle and the menisci appearance \cite{pagonabarraga-finger} should not
influence the overall solution.
 \item The free-energy model examined in the work is single density, two
viscosities model. The classical Bretherton problem is a free-interface problem
where the film thickness is established by itself. The viscosity ratio should
be large, as indicated in \cite{pagonabarraga-fingers}, however no studies
address this issue. \todo{what's the conclusion of the impact of this issue to our paper then?}
  \item The Bretherton problem addresses bubble flow driven by a pressure
difference, not a body force.  It is however relatively difficult to impose a
pressure difference in the framework of the binary liquid model, as two
simultaneous boundary conditions with the associated gradients and laplacians are required.
This kind of pressure boundary conditions for the binary liquid model is currently
under development, and preliminary results show that the differences between
pressure difference flows and flows driven by a body force are small as far as the film
deposition thickness is concerned.  In this work we limit ourselves to the study
of body force driven flows, because of their simplicity and better numerical stability.
 \item Most simulations are done for a circular channel,
which is quite difficult to address in terms of lattice Boltzmann free energy
framework where one needs to calculate Laplacians near the boundaries.
Although one can apply certain finite difference stencils
\cite{arnold-boundary,hunt-boundary} developed specifically for curved boundaries,
the analysis of error introduced by the boundary conditions becomes an issue.
Our benchmark is therefore done is for an axis-symmetric case. \todo{since this paper
is about a 2d flow, which we already said in the intro, what is the point of discussing
curve boundaries which are only an issue in 3d?}
\end{itemize}
Given all the concerns and challenges, the suggested lattice Boltzmann framework
is a two-dimensional flow driven by a body force.  An analytical
solution to this problem is to the best of our knowledge not yet known.
Though some authors \cite{sehgal-microchannel} use the gravity driven model to
validate the Bretherton film thickness, the latter should be used with caution
since the velocity of the bubble is different from the fluid velocity and the
bubble shape
is different in the front and rear menisci. \citet{wong-films} indicated that the
film width for polygonal capillaries
changes from $\Ca^{1/2}$ on distances $x\propto \Ca^{-1}$ to $\Ca^{2/3}$ for
distances $x \gg \Ca^{-1}$. Therefore, for a bubble to have a formed shape in
the  front meniscus we chose the bubble length to be $5$ channel heights. The
channel length is taken  as $3$ bubble lengths (or $15$ channel heights)
to minimize the influence of one bubble on another,
because of the periodicity of boundary conditions.
Figure \ref{fig:benchmark:sketch} represents a sketch of the geometry
used in the benchmark.
\begin{figure}
%\includegraphics[width=0.97\textwidth]{}
\caption{{\color{red} Ask Sasha to help here. The benchmark sketch.}
\label{fig:benchmark:sketch}}
\end{figure}

\section{Results}
\subsection{The nondimensionalization and initialization procedure}
We now intrduce the procedure of nondimensionalization, as well as
the initialization techniques we used to be able to predict parameters necessary for
simulation. The capillary number governs the interface thickness and is defined as:
\begin{equation}
Ca=\frac{U_{\mathrm{bubble}} \mu_{\mathrm{liq}}}{\gamma},
\end{equation}
where $\mu_{\mathrm{liq}}$ is the liquid viscosity, and $\gamma$ is the surface tension. The
bubble velocity $U_{\mathrm{bubble}}$ is taken in the center of the bubble where the film
thickness is examined.
While parameters such as $\mu_{\mathrm{liq}}$, $\gamma$ and the capillary number $\Ca$ are given
or defined through the model parameters, the velocity $U_{\mathrm{bubble}}$ is not.
In the simulation we approximately match the velocity with the
Poiseuille profile center velocity, which is given by the equation:
\begin{equation}
\begin{aligned}
U_{\mathrm{bubble}} \approx U_{\mathrm{liq}}=\frac{H^2}{8
\mu_{\mathrm{liq}}}\frac{\mathrm{d}P}{\mathrm{d}x}\\
U_{\mathrm{bubble}}\approx
\frac{{N_y}^2}{\frac{8}{6}(\tau_{\mathrm{liq}}-\frac{1}{2})}\frac{\mathrm{d}P}{\mathrm{d}
x } ,
\end{aligned}
\end{equation}
By taking the Poiseuille profile one can estimate the body force through the
capillary number and grid resolution:
\begin{equation}
\label{poiseuille:velocity:center}
\begin{aligned}
&\frac{\mathrm{d}P}{\mathrm{d}x}=\frac{8}{{N_y}^2}\gamma Ca.
\end{aligned}
\end{equation}
In practice, we iterate adjusting the body force value until both the desired velocity and the
capillary numbe are achieved.

After one calculates the parameters necessary to run a simulation, one needs to
initialize the macroscopic fields and particle populations.
The velocity is initialized with $0$ everywhere. Populations are initialized
with the equilibrium values for the binary liquid model, including all the phase
gradients and Laplacians. The liquid phase is initialized with $\phi = 1$ and the
gas phase is initialized with $\phi = -1$. Though the initialization film thickness
is not involved in the final result for the front meniscus (see Fig.
\ref{fig:different:initialization:widths}) we keep the initialized film
thickness as close as possible to the already obtained numerical film
thickness values \cite{giavedoni-numerical}. This is done to minimize
the simulation time, as starting far away from the steady-state requies
additional time for the relaxation.
\begin{figure}
\includegraphics*[bb=-20 360 620 430,width=0.97\textwidth]
{Figures/Init/initbegin12.eps}
\includegraphics*[bb=-20 360 620 430,
width=0.97\textwidth]{Figures/Init/initbegin16.eps}
\includegraphics*[bb=-20 360 620 430,
width=0.97\textwidth]{Figures/Init/initbegin20.eps}
\includegraphics*[bb=-20 360 620
430,width=0.97\textwidth]{Figures/Init/initfinish12.eps}\\
\includegraphics*[bb=-20 360 620
430,width=0.97\textwidth]{Figures/Init/initfinish16.eps}\\
\includegraphics*[bb=-20 360 620
430,width=0.97\textwidth]{Figures/Init/initfinish20.eps}\\
\caption{The phase plots for different bubble widths as
$H_{\mathrm{eff}}-12$,$H_{\mathrm{eff}}-16$ and
$H_{\mathrm{eff}}-20$. The grid for the simulation is $102 \times 1501$. The phase
profiles were obtained after $2\cdot10^5$ time steps. One can see that even if the
system is initialized with different bubble volumes, the bubbles always relaxate
to a shape where the film thickness is conserved. For the given parameters the
interface thickness is $0.0615$, $0.0611$, $0.0605$.
\label{fig:different:initialization:widths}}
\end{figure}

Here, two different examples are presented that can quantitatively show the
initialization procedure for quite different capillary numbers. In these
examples we use $k=0.04$ and $A=0.04$ as optimal binary liquid model parameters.
That explicitly implies the surface tension to be $\sqrt{\frac{8 k
A}{9}}=0.0377$ and the characteristic length of the interface
$5\sqrt{\frac{k}{A}}=5$ lattice boltzmann units. For the sake of simplicity and
stability the relaxation times for liquid and gas phases are taken as
$\tau_{\mathrm{liq}}=2.5$ and $\tau_{\mathrm{gas}}=0.7$, which corresponds
to a gas-liquid viscosity ratio of $10$.
\begin{description}
 \item[I \bm{$Ca=0.005$}]
  The film width for such small Capillary numbers \cite{bretherton} is
given by the equation:
  \begin{equation}
  \frac{h_{\infty}}{H}=1.3375 \Ca^{2/3}=0.03899,
  \end{equation}
  where we assume that $2 H$ or $N_y$ is the channel width in lattice units.

\todo{Therefore one needs to resolve the interface which is around $2$ percent of
the  whole channel width with, for instance $8$ lattice units, where $5$ lattice
units occupies the interface width for given. Note that the interface occupies
$62.5$ percent of the interface. -- these two sentences need to be rewritten for clarity}  Therefore the channel
  width is $400$ lattice units.
